MLH1 Results Outline

### 2020-02-18 doc/Results.rmd
finer editing for the begining MLH1 section

@@

```{r table.for.plot, echo=FALSE}
#why do I add in a different table?

@@

```{r main.fig1s, echo=FALSE, include=FALSE}

strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

@@

main.table <- add_subsp(main.table)

@@

In order to estimate the genome wide recombination rates across our panel of `r length(unique(main.table$strain))` strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)`.

The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, and _M. m. domesticus$^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^PWD^_, _M. m. musculus^SKIVE^_, and _M. m. musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^ MSM^_ and _M. m. molossinus^MOLF^_ respectively.

While in two of the subspecies, the means for males show a greater range across means, notable in the mmmolossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^MSM^_ and  _M. m. molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^PWD^_, _M. m. musculus^SKIVE^_, _M. m. musculus^KAZ^_, _M. m. musculus^TOM^_, _M. m. musculus^AST^_, and _M. m. musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, _M. m. domesticus^LEW^_ and _M. m. domesticus^PERC^_ respectively.

<!-- mouse level means-->

<!--(the figure is made from pooling MLH1 cells)
the male

- the ranges of mouse means are displayed in Sup table X

- How the means were calculated

-The male gwRR is adjusted for the PAR
-The means for each strains are report in Table 1

-Normality was assessed for the MLH1 counts (include in Supplement histograms)

, -->

@@

In order to test the effects of subspecies, sex and strain on the mean and variance of MLH1 counts per cell, we fit a data set of `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages for MLH1 foci per cell. We constructed the model to set subspecies, sex and their interaction as fixed effects, which strain was coded as a random effect. The predicted coeffecients and pvalues are reporter below in table X.

$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

<!--
This is the break down of mice used within the Mixed model

@@

add number of cells

`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`

 $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

In order to understand the variance within an evolutionary framework, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse level MLH1 foci per cell means as the dependant variable with a mixed model.

@@

summary(MLH1.MM.M1)$coefficients

@@

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)

Because all of the mixed model effects had significant effects, (p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r  round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect), we followed up with two general linear models to investigate the specific strain effects.

 $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$

and

$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$

@@

In the first glm, the fixed subspecies effects were not significant, so we only report for values for the second model which is limited to sex and strain fixed effects.

@@

round(summ.strain.fixed$coefficients[,c(1,4)],5)

@@

In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _M. m. musculus^PWD^_ and _M. m. molossinus^MSM^_  respectively. These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.

Additionally, _M. m. domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _M. m. domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.

The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant. The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.

# Within Mouse Variance in CO Count per Cell

In order to more fully discribe the distributions we examined the variance in addition to the mean applying the same mixed models and linear models, with _within mouse variance_ of MLH1 counts per cell as the dependant variable.

limitations: observations at the mouse level, decreased power to estiamate the variance --

setup

-quality / technical error

- variance and cv

The mixed models have most significant variables,

M3.glm for variance was less significant (0.044181 ),

the Q12 models in general were more significant

cV's were more significant than var

$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$

@@

LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
LmerMLH1_VAR_results

@@

LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
LmerMLH1_CV_results

@@

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
LmerMLH1_M1_Q12_VAR_results

@@

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)

LmerMLH1_M1_Q12_VAR_results
```

These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12

```{r cv.plots, echo=FALSE, eval=FALSE}

cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +
  ggtitle("cV of within mouse variance, all mice")
cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
cV.Hist.total.mice

#DF.HetC.MixedModel.HQ
cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  + 
  ggtitle("cV of within mouse variance, 10 cell min")
cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
cV.Hist
```

@@

We followed up the mixed models with glms. The coefficients are listed below for the following glms: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12

```{r M3.MLH1.var, echo=FALSE}

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )

MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)

Across all four data sets, the sex effect has the most significant p value. The highest p value is for the sex effect for VAR and Cv in the full models. This might be due to extra noise from technical sources.

<!--
var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***

For both the mixed model and linear model, sex was the most signifcant effect (p= `r LmerMLH1_VAR_results$sex.results` p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` (for the high quality dataset)  ) respectively.

The measures for variance are more supciptibal to thechnical error effects (noise). For this reasons, we analysze the models by restricting the data to a subset cells which were scored as highest quality. Sex was also the most significant effect (p)

<!--
Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across

Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.

<!--

@@

## Female Specific Analysis

In order to more finely discirbe the sex specific patterns, we foucsed on data sets including single sexes.

@@

                        total.cells  = sum(Ncells),

                        mouse.mean_co = mean(mean_co),

@@

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
F.strain.table

@@

For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
To analyze the female specific dataset,

In the dom  subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M. m. domesticus$^WSB^_ ,

@@

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _M. m. domesticus$^LEW^_

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _M. m. musculus$^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _M. m. musculus$^KAZ^_

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _M. m. molossinus$^MSM^_

Two glm models were run to test the effects of strain and subspecies on mean MLH1 coount per cell per mouse. The coefficients for the following model are below.

@@

summry.female.strain.effects <- summary(glm.female.strain.subsp)
round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)

@@

round(lm.female.strain_sum$coefficients[,c(1,4)],5)

@@

For mean CO count, _M. m. domesticus^G^_  strain effects in both models (p = )
_M. m. musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).

@@

_M. m. domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means. _M. m. domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _M. m. musculus^MSM^_ is `r mean.MSM.female / mean.all.other`. These three will be designated as 'moderate high rec' strains for later sex specific analysis.

@@

#round( #round isn't working for this sumary
lmVAR.female.strain_sum$coefficients[,c(1,4)]

@@

#round(
lm.m1.Q12_female_sum$coefficients[,c(1,4)]
#LEW not sig

@@

lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig

The variance across mouse means for each strain is also variable across strains. With _M. m. domesticus$^LEW^_ having the most significant strain effect in the second model (p= ).

## Male Specific Analysis

The male specific patterns were also investigated.

@@

@@

Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))

@@

#female strain table
#female_mouse.avs_for.analysis
M.strain.table <- ddply(just.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  = sum(Ncells),
                        mean_co =as.numeric(format(round(mean(mean_co), 4 ), nsmall=3) ),

@@

                        mean.mouse_sd   = mean(sd),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
M.strain.table
#same issue as the female table .. ugh
```

@@

 _M. m. domesticus^LEW^_

@@

round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

@@

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)

#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig

lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant

@@

lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS

lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS

The coefficients for the variance models are listed above

The strain average for  _M. m. musculus^PWD^_ , _M. m. molossinus^MSM^_ and  _M. m. musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

@@

```{r Q1_short_biv_lmer, echo=FALSE}

lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

#skive strin very slight sign
#male highly sig
#skive * male slightly sig

@@

```{r Q1.strain.specific.junk, include=FALSE}

#Dom.all
#Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.Dom)

#Dom.1CO
#Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.1CO.Dom)

>>>>

female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")

@@

```{r table.for.plot, echo=FALSE, warning=FALSE}
#why do I add in a different table?

@@

```{r main.fig1s, echo=FALSE, include=FALSE}
strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

@@

main.table <- add_subsp(main.table)
main.table <- add_species(main.table)

@@

In order to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` _M. m. musculus_ strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)` oocytes.

(mouse means were calculated and statisics in Table 2)

Figure 1(A) shows the --

--notes on error /

-- connections to previous measures in the literature

<!--
The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively. 
-->

While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.

<!--(the figure is made from pooling MLH1 cells)
the male 
- the ranges of mouse means are displayed in Sup table X
- How the means were calculated
-The male gwRR is adjusted for the PAR
-The means for each strains are report in Table 1
-Normality was assessed for the MLH1 counts (include in Supplement histograms)
-->

@@

In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.  (*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )

<!--
$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-->
<!--
This is the break down of mice used within the Mixed model

@@

add number of cells
`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
 $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

@@

#summary(MLH1.MM.M1)$coefficients

@@

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)
```

Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.

All of the effects from the mixed model significant, (LRT; p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect).

<!--
 $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
and 
$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$  -->

@@

We followed up the mixed model effects with general linear models. The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions. (*what are the points and differences? )*

@@

#round(summ.strain.fixed$coefficients[,c(1,4)],5)

@@

In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively. These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.

Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.

## Within Mouse Variance in CO Count per Cell

In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).

@@

LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)

@@

LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)

@@

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)

@@

LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
```
<!--
$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$-->
```{r M3.MLH1.var, echo=FALSE}

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)

#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```
<!--
These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12

For both the mixed model and linear model, sex was the most significant effect while the subspecies, interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively.

*notes on caveats for calculating variance -- lower power

Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e. background noise). For this reasons, we replicated the model analysis from data sets in the subset of cells with higher quality scores compared to the full data set. (these analysis/ models replicated the results of the full data set (sex being the most significant 
(LTR; p = `r LmerMLH1_M1_Q12_VAR_results$sex.results`, and glm; p = `r  MLH1var_M3_Q12_sum$coefficients[2,4]` )

**female strongest effect, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.

limitations: observations at the mouse level, decreased power to estiamate the variance --
setup
-quality / technical error
- variance and cv
The mixed models have most significant variables, 
M3.glm for variance was less significant (0.044181 ), 
the Q12 models in general were more significant
cV's were more significant than var

cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***

$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
<!--
Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across

Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
<!--

@@

@@

# Female Specific Analysis

In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.

@@

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

@@

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
F.strain.table

@@

For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
To analyze the female specific data set,

In the _domesticus_  subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,

@@

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _domesticus^G^_

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _ musculus^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_

Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.

@@

summry.female.strain.effects <- summary(glm.female.strain.subsp)
#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)

@@

#round(lm.female.strain_sum$coefficients[,c(1,4)],5)

@@

For mean CO count, _domesticus^G^_  strain effects in both models (p = )
_musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).

@@

_domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means. _ domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _musculus^MSM^_ is `r mean.MSM.female / mean.all.other`. These three will be designated as 'moderate high rec' strains for later sex specific analysis.

@@

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]

@@

#round(
#lm.m1.Q12_female_sum$coefficients[,c(1,4)]
#LEW not sig

@@

#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig

The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).

# Male Specific Analysis

The male specific patterns were also investigated. (what's a better motivation for this section)

@@

#fomo.male_mouse.av

@@

Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
```

@@

M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

@@

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
M.strain.table
#same issue as the female table .. ugh
```

@@

 _domesticus^LEW^_

@@

#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

@@

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)

#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig

lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant

@@

lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS

lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS
```

The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

@@

```{r Q1_short_biv_lmer, echo=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

#skive strin very slight sign
#male highly sig
#skive * male slightly sig
```

@@
